Adaptive Waveform Inversion for Transmitted Wave Data

### 2024-11-05 repro/papers/awi/paper.tex
interim

\section{Introduction}
Full waveform inversion (FWI) refers to the inference of mechanical
parameter distribution within a material object  (the earth, a human
body, a manufactured artifact,...) from remote measurements of waves
propagating through it. These mechanical parameters appear as
coefficients in systems of partial differential equations governing
wave propagation, and numerical solutions of these wave equations can
be used to predict experimental data {\em ab initio}. FWI consists in
updating the parameters to minimize a numerical measure of  the difference between computationally
predicted wave data and actual observed
data. If wave velocities or their parametric equivalents are amongst the parameters
to be estimated, such numerical measures tend to oscillate, due to:
the excessive sensitivity of predicted waveforms to changes in wave velocity.
%especially for the most commonly used square-integral measure ($L^2$
%norm).
These oscillations cause iterative local optimization algorithms (relatives of
Newton;'s method, the only practical algorithmic choices) to stagnate
far from good data fit \cite[]{VirieuxOperto:09}, unless the initial
parameter estimates are of very high quality.

Adaptive Waveform
Inversion \cite[]{Warner:16} (AWI) is a modification of FWI, designed to overcome this
``cycle-skipping'' pathology. AWI has been successfully applied to
industry-scale seismic data
\cite[]{GuaschWarnerRavaut:GEO19,Warneretal:SEG21} and shows promise
for ultrasonic imaging of the the human brain
\cite[]{Guaschetal:NPJDM20}. These and other works discuss heuristic justifications for AWI
and many details of implementation. However, a clear explanation for its
effectiveness is so far missing from the literature.

This paper supplies such an explanation for a special case of AWI, the recovery of smooth acoustic
medium parameters from data generated at localized energy sources,
propagated through a fluid, and recorded at receiver locations remote
from the sources. I show that the AWI objective function is an
asymptotic (in time frequency) approximation to the mean-square
time-of-travel misfit, provided that arrival times of transmitted
waves are single, that is, unique rays of geometric asymptotics
connect sources and receivers. The error in this approximation is
proportional to the source mean-square wavelength. Thus AWI and
minimization of mean-square travel time error (travel time tomography)
lead to similar material parameter estimates. Travel time tomography
appears to be much less sensitive to choice of initial model than is
FWI, and in fact is commonly used to construct an initial model for
FWI \cite[]{Bordingetal:87,SirguePratt:04,VirieuxOperto:09}. At least
in the special case treated here, AWI implicitly combines travel time
tomography with FWI.

The reader will
note that a normalization component of the AWI objective function
definition is important in reaching this conclusion. The connection
between objective normalization and the asymptotic equivalence to
travel time inversion for transmission data has nothing to do with the
reasons for normalizing the objective previoiusly advanced in the
literature \cite[]{Warner:16,LiAlkhalifah:21}. Another similar modification of
FWI, Matched Source Waveform Inversion (MSWI)
\cite[]{HuangSymes2015SEG,HuangSymes:Geo17} lacks the AWI
normalization, and has a looser relation to travel time inversion. The
arguments presented here for AWI (and MSWI) are refinements of those
given by \cite{HuangSymes:Geo17}. which were in turn based on ideas
presented in Song's thesis \cite[]{Song:94c}.

The next section establishes the class of problems to be studied and notations
for their description, defines the AWI and MSWI objective functions,
and states the main results of this paper. in the following section, I present the
asymptotic analysis leading to the main results. In
effect, AWI regularizes or relaxes the FWI problem. For many inverse
problems, regularization has been accomplished via additive penalty
functions. In section four, I explain the relation between AWI and a
penalty formulation of the wave inverse problem. The close relation
between AWI objective and travel time misfit is limited in scope: in
the fifth section, I show that it does not hold in general for
transmitted wave data in which multiple rays connect source and
receiver.

@@

The parameter to be determined in the inverse problem studied here is
the bulk modulus $\kappa$. It is presumed is bounded and smooth:
\begin{equation}
  \label{eqn:coeff}
  \log \kappa \in C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)
\end{equation}
Under these hypotheses, the symmetric hyperbolic system \ref{eqn:awe}
has a unique distribution solution (for example \cite{Lax:PDENotes},
Theorem 7.4). The Propagation of
Singularities Theorem (\cite{Tay:81}, Theorem VI.2.1) %and
%the pull-back property of distributions (\cite{Dui:95}, Proposition
%1.3.3)
imply that $p(\cdot,\cdot;\bx_s)$ is smooth away from $\{\bx_s\}
\times \bR$, so that the traces $p(\bx_r, \cdot;\bx_s)$ on
$\{(\bx_r,t): t \in \bR\}$ are well-defined.

Define the {\em modeling operator}
\begin{equation}
  \label{eqn:modop}
  F: exp(M)  \rightarrow \Pi_{X_{sr}} {\cal D}'(\bR)
\end{equation}
by
\begin{equation}
  \label{eqn:modopdef}
  F[\kappa]= p|_{X_{sr} \times \bR}
\end{equation}

@@

The analysis to follow requires two further constraints on the domain
$M \subset C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)$ of $F$. The first is
the {\em Exponential Decay} assumption:
\begin{quote}
  For any $n > 0$, $f \in {\cal E}'(\bR)$,
  there exist $K >
  0, T \in \bR,$ and $\delta>0$ so
  that for any $k$ with $0 \le k \le n$, $(\bx_s,\bx_r) \in
 X_{sr},$ and any $\log \kappa \in M$, the solution $(p,\bv)$ of \ref{eqn:awe} is smooth
  for $t>T$ and satisfies
  \[
    |\partial_t^k p(\bx_r,t;\bx_s)| \le K e^{-\delta t}, \, t > T.
  \]
\end{quote}
SInce $X_{sr}$ is bounded, the property described in this assumption
is a special case of {\em local energy decay}, a much studied topic. The {\em
  non-trapping} hypothesis is sufficient to ensure local energy decay:
all rays of geometric asymptotics (that is, geodesics for the Riemannian metric associated
to the system \ref{eqn:awe}) escape to infinity (\cite{Hristova:09},
Theorem 2, \cite{EgorovShubin}, Theorem 2.104, p. 192). That is,
exponential decay follows from a characteristic of $\kappa$.

The second assumption explicitly involves a constraint on ray geometry
(therefore on $\kappa$). A {\em normal neighborhood} of
$\bx_s \in \bR^3$ is an open subset $\Omega$ of $\bR^3$ 
within which every point is connected to $\bx_s$ by  a unique ray  (\cite{Friedlander:75},
p. 16). Therefore, to every point $\bx \in \Omega$ is associated a
unique time of travel (geodesic distance) $\tau(\bx_s,\bx)$ from $\bx_s$ to $\bx$.
The {\em Single Arrival} assumption is
\begin{quote}
  $\Omega\subset \bR^3$ 
and $M \subset C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)$ are bounded open sets satisfying 
\begin{itemize}
  \item[G1. ] $X_{sr} \subset \Omega \times \Omega$;
  \item[G2. ] for every source point $\bx_s\in X_{s}$ and $\log \kappa \in M$, $\Omega$ is a
    normal neighborhood of $\bx_s$.
  \item[G3. ] for every source point $\bx_s \in X_{s};$
    $\log \kappa \in M$, ray ${\bf X}:\bR^+\rightarrow \bR^3$
    with ${\bf X}(0)=\bx_s$ and any $t>0$, either $\{{\bf X}(s): 0\le s \le t\} \subset \Omega$
    or $\{{\bf X}(s): t < s < \infty\} \subset \bR^3 \setminus \Omega$.
  \end{itemize}
\end{quote}
Assertion G3 is the {\em no-return} assumption: once a ray leaves
$\Omega$, its subsequent intersection with $\Omega$ is null.

Thanks to exponential decay, each trace $p(\bx_r,\cdot;\bx_s)$ is in $L^{p}(\bR)$ for $p
\ge 1$.
Denote by $L^p(X_{sr} \times \bR)$ the product space $l^p(X_{sr},L^p(\bR))$
with the usuall product norm.

@@

AWI modifies the least-squares FWI problem by interposing a convolution operator or
{\em filter} $u \in L^2(X_{sr} \times \bR)$. That is, the residual
data $F[\kappa]-d$ is replaced by the filtered residual data $F[m]*u-d$.
Since $F[\kappa] \in L^1(X_{sr}\times \bR)$, the filtererd residual
makes sense.
The filter $u$ is intended to shift localized wave energy in $F[m]$ to
match that present in $d$. If on the other hand $u(\bx_r,t;\bx_s) =
\delta(t)$ for each $(\bx_x,\bx_r) \in X_{sr}$, then $tu = 0$. This
observation motivates the definition of the AWI objective function
\begin{equation}
  \label{eqn:awidefmain}
  J_{\rm AWI,\sigma}[\kappa;d] = \sum_{(\bx_s,\bx_r) \in X_{sr}}
  \frac{\|Tu_{\sigma}[\kappa;d](\bx_r,\cdot;\bx_s)\|^2}{\|u_{\sigma}[\kappa;d](\bx_r,\cdot;\bx_s)\|^2}.
\end{equation}
in which $Tu(\bx_r,t;\bx_s) = tu(\bx_r,t;\bx_s)$ and
\begin{equation}
  \label{eqn:reg}
  u_{\sigma}[\kappa;d] = \mbox{argmin}_{u \in L^2(X_{sr} \times \bR)} (\|F[\kappa]*u-d\|^2 + \sigma\|u\|^2)
\end{equation}
The objective function for the closely related MSWI approach,
mentioned in the introduction, is  
\begin{equation}
  \label{eqn:mswidefmain}
  J_{\rm MSWI,\sigma}[\kappa;d] = \sum_{(\bx_s,\bx_r) \in X_{sr}}
  \|Tu_{\sigma}[\kappa;d](\bx_r,\cdot;\bx_s)\|^2.
\end{equation}
Note that while the Lax-Milgram Theorem (\cite{Yosida}, section
III.7) ensures that $u_{\sigma}[\kappa;d] \in L^2(X_{sr}\times \bR)$,
it is not immediiately obvious that $Tu_{\sigma}[\kappa;d] \in L^2(X_{sr}\times \bR)$.

Asymptotic analysis of AWI and MSWI is based on a family of source
wavelet: for $\lambda > 0$,
\begin{equation}
  \fl(t)= \lambda^{\frac{1}{2}}f_1\left(\frac{t}{\lambda}\right).
\end{equation}
The relation between the dimensionless scale parameter $\lambda$ and
RMS wavelength will be explained below. Correspondingly, denote the
modeling operator with $f=\fl$ by $\Fl$.

@@

\begin{theorem}
  \label{thm:main}
  Assume that the Exponential Decay and Single Arrival assumptions
  hold for $X_{sr} \subset \Omega \times \Omega, M \subset C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)$.
  Select $r >0$ with units of
  pressure$^2 \times$time$^2$ and $\log \kappa^* \in M$. For $0 <
  \lambda $, set
  $\dl=\Fl[\kappa^*]$ (``noise-free data''), and  define $\ul[\kappa,\dl]$
  to be the solution of 
  equation \ref{eqn:reg} with  $\sigma=r\lambda$, $F=\Fl$ and $d=\dl$..
  There exist $0 < C_* \le C^*$, $0 < C$, and $W: M \times X_{sr}
  \rightarrow \bR^+$ with $C_* \le
  W \le C^*$ so that for
  $\log \kappa \in M$, $(\bx_s,\bx_r) \in X_{sr}$, $Tu_{\sigma}[\kappa;\dl](\bx_r,\cdot;\bx_s) \in L^2(\bR)$, and 
  \begin{equation}
    \label{eqn:mswitt}
    |\|T\ul[\kappa;\dl](\bx_r,\cdot;\bx_s)\|- \lambda^{-\frac{1}{2}} (W[\kappa^*,\kappa](\bx_s,\bx_r))^{\frac{1}{2}}|\tau[\kappa](\bx_s,\bx_r)-\tau[\kappa^*](\bx_s,\bx_r)|| \le C
  \lambda^{\frac{1}{2}}
\end{equation}
and for sufficiently small $\lambda$, 
\begin{equation}
  \label{eqn:awitt}
  \left|\frac{\|T\ul[\kappa;\dl](\bx_r,\cdot;\bx_s)\|}{\|\ul[\kappa,\dl](\bx_r,\cdot;\bx_s)\|}-
    |\tau[\kappa](\bx_s,\bx_r)-\tau[\kappa^*](\bx_s,\bx_r)|\right| \le C \lambda
  \end{equation}
\end{theorem}

Summation over the source-receiver positions in $X_{sr}$ yields
\begin{cor}
  \label{thm:cor1}
  In addition to the assumptions of Theorem \ref{thm:main}, define $J_{\rm
    AWI,\sigma}$ and $J_{\rm MSWI,\sigma}$ by equations
  \ref{eqn:awidefmain}, \ref{eqn:mswidefmain} respectively. Then
\begin{equation}
\label{eqn:cor1mswi}   
\lim_{\lambda \rightarrow 0} \lambda J_{\rm MSWI,r\lambda}[\kappa;\dl]
= \sum_{(\bx_s,\bx_r) \in X_{sr}} W[\kappa^*,\kappa^*]^2(\bx_s,\bx_r) |\tau[m](\bx_s,\bx_r) - \tau[m^*](\bx_s,\bx_r)|^2.
\end{equation}
and
\begin{equation}
\label{eqn:cor1awi}   
\lim_{\lambda \rightarrow 0} J_{\rm AWI,r\lambda}[\kappa;\dl]
= \sum_{(\bx_s,\bx_r) \in X_{sr}}  |\tau[m](\bx_s,\bx_r) - \tau[m^*](\bx_s,\bx_r)|^2.
\end{equation}
\end{cor}

\subsection{Hadamard's Decomposition}
The Hadamard decomposition of the acoustic Green's function \cite[]{Friedlander:75,Qian:JCP24} provides a
basis for a norm-based formulation of FWI. As presented by
\cite{Friedlander:75}, for instance, this construction applies to
second-order wave equations, but the system \ref{eqn:awe} is
equivalent fo a second-order problem: differentiate the first equation
with respect to $t$, eliminate $\bv$ via the second equation, and
rearrange to obtain
\begin{equation}
  \label{eqn:awe2}
  \frac{\rho}{\kappa}\frac{\partial^2 p}{\partial t^2} - \nabla^2 p -
  \nabla \log \rho \cdot \nabla p = \rho \frac{\partial f}{\partial t}
  \delta(\bx-\bx_s), \nonumber \\
  p=0, \, t \ll 0.
\end{equation}
If $(p,\bv)$ is a distribution solutions of the system
\ref{eqn:awe} then $p$ is a solution of equation
\ref{eqn:awe2}, as follows by differentiating the first equation of
\ref{eqn:awe} with respect to $t$ and eliminating $\bv$ via the second
equation. Conversely, if $p$ is a solution of the second-order initial
value problem\ref{eqn:awe2}, then under suitable conditions (including
those to be imposed below),
\[
  \bv = \beta \int_{-\infty}^t p
\]
defines a distribution, and $(p,\bv)$ solves the system \ref{eqn:awe}.

Hadamard's construction is local, in both space-time and in the set of
acoustic models. A {\em normal neighborhood} of
$\bx_s \in \bR^3$ is an open subset of $\bR^3$ 
within which every point is connected to $\bx_s$ by  a unique ray of
geometric acoustics (a geodesic for the Riemannian metric associated
to the equation \ref{eqn:awe2}) (\cite{Friedlander:75},
p. 16).

Throughtout the following discussion, $\Omega\subset \bR^3$ 
and $M \subset C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)$ are bounded open sets satisfying 
\begin{itemize}
  \item[G1. ] $X_{sr} \subset \Omega \times \Omega$;
  \item[G2. ] for every source point $\bx_s\in X_{s}$ and $\log \kappa \in M$, $\Omega$ is a
    normal neighborhood of $\bx_s$.
  \item[G3. ] for every source point $\bx_s \in X_{s};$
    $\log \kappa \in M$, ray ${\bf X}:\bR^+\rightarrow \bR^3$
    with ${\bf X}(0)=\bx_s$ and any $t>0$, either $\{{\bf X}(s): 0\le s \le t\} \subset \Omega$
    or $\{{\bf X}(s): t < s < \infty\} \subset \bR^3 \setminus \Omega$.
  \end{itemize}
Condition G3 (the ``no return'' condition) means that once a ray
leaves $\Omega$, it does not return to $\Omega$. Conditions G1 - G3
hold for any convex open set $\Omega$ if $\kappa$ is constant, and for small perturbations and some
large perturbations of constant $\kappa$: these are open conditions in
the topology of $M$.

Under conditions G1 - G4, the construction described by
\cite{Friedlander:75} produces for each $m \in M$ functions $\tau[\kappa]$ (travel time), $a[\kappa]$
(geometric amplitude) on $X_s \times \Omega$ and $b[\kappa]$
(remainder kernel) on $X_s \times \Omega \times \bR$, satisfying
\begin{itemize}
\item[H1. ] $\tau[\kappa], a[\kappa],$ and $b[\kappa]$ are smooth off the diagonal,
  that is,
  (i) for each $\bx_s \in X_s$, $\tau[\kappa](\bx_s,\cdot) \in C^{\infty}(\Omega \setminus
  \{\bx_s\})$, and similarly for $a[\kappa],b[\kappa]$; $\tau$
  satisfies the eikonal equation
  \[
    \sqrt{\frac{\kappa}{\rho}} |\nabla \tau(\bx_s,\cdot)| = 1 \mbox{ in } \Omega \setminus
    \{\bx_s\}, \, \lim_{\bx \rightarrow \bx_s} \tau(\bx_s,\bx) = 0.
  \]
\item[H2. ] there exists $A>0$ so that for any $\log \kappa \in M$
  and $(\bx_s,\bx_r) \in X_{sr}$,
  $|\log a[\kappa](\bx_s,\bx_r)| < A$;
\item[H3. ] for any $k \ge 0$, there exists $B \ge 0,\delta > 0$ so
  that for any $(\bx_s,\bx_r) \in X_{sr}$, $\kappa \in M$, 
\begin{equation}
  |\partial^k_t b[\kappa](\bx_r,t;\bx_r,s)| \le B e^{-\delta t}
  \label{eqn:best}
\end{equation}
\end{itemize}
(The exponential decay estimate in item H3 follows from hypothesis G4.)

@@

>>>>

\section{Introduction}
Full waveform inversion (FWI) refers to the inference of mechanical
parameter distribution within a material object  (the earth, a human
body, a manufactured artifact,...) from remote measurements of waves
propagating through it. These mechanical parameters appear as
coefficients in systems of partial differential equations governing
wave propagation, and numerical solutions of these wave equations can
be used to predict experimental data {\em ab initio}. FWI consists in
updating the parameters to minimize a numerical measure of  the difference between computationally
predicted wave data and actual observed
data. This data is typicallyl oscillatory. If wave velocities or their parametric equivalents are amongst the parameters
to be estimated, then updates cause shifts in the arrival times of
data oscillations, which in turn causes commonly used numerical error measures
to oscillate. This ``cycle skipping'' phenomenon may cause iterative local optimization algorithms (relatives of
Newton;'s method, the only practical algorithmic choices) to stagnate
far from good data fit, unless the initial
parameter estimates accurately predict the arrival times of waves \cite[]{GauTarVir:86,VirieuxOperto:09,HuetalIEEE:18}.

Adaptive Waveform
Inversion \cite[]{Warner:16} (AWI) is a modification of FWI, designed to overcome the
cycle-skipping pathology. AWI has been successfully applied to
industry-scale seismic data
\cite[]{GuaschWarnerRavaut:GEO19,Warneretal:SEG21} and shows promise
for ultrasonic imaging of the the human brain
\cite[]{Guaschetal:NPJDM20}. These and other works discuss heuristic justifications for AWI
and many details of implementation. However, a clear explanation for its
effectiveness is so far missing from the literature.

This paper supplies such an explanation for a special case of AWI, the
recovery of smooth acoustic medium parameters from data generated at
localized energy sources, propagated through a fluid, and recorded at
receiver locations remote from the sources (transmission data). I show that the AWI
objective function is an asymptotic (in time frequency) approximation
to the mean-square time-of-travel misfit, provided that arrival times
of transmitted waves are single. That is, a unique ray of geometric
asymptotics connects each sources-receiver pair, as is the case if the associated
Riemannian metric is simple in a domain containing the sources and
receivers \cite[]{PestovUhlmann:05,StefanovUhlmann:05}. The error in
this approximation is proportional to the source mean-square
wavelength. I also show that this relation fails if arrival times have
higher multiplicity. Travel time tomography (minimization of
mean-square travel time error) appears to be much less sensitive to
choice of initial model than is FWI. In fact, it is commonly used to
construct an initial model for FWI
\cite[]{Bordingetal:87,SirguePratt:04,VirieuxOperto:09}. At least for
acoustic transmission data with single arrivals, AWI implicitly
combines travel time tomography with FWI.

The arguments presented here for AWI are refinements of those given by
\cite{HuangSymes:Geo17} for Matched Source Waveform Inversion (MSWI),
a similar modification of FWI
\cite[]{HuangSymes2015SEG,HuangSymes:Geo17}. which were in turn based
on ideas presented in Song's thesis \cite[]{Song:94c}. Since it
involves no additional effort, I state results for MSWI in parallel
with those for AWI.

The next section establishes the class of problems to be studied and notations
for their description, defines the AWI and MSWI objective functions,
and states the main results of this paper. in the following section, I present the
asymptotic analysis leading to the main results. In
effect, AWI regularizes or relaxes the FWI problem. For many inverse
problems, regularization has been accomplished via additive penalty
functions. In section four, I explain the relation between AWI, MSWI, and 
penalty formulations of the wave inverse problem. The close relation
between AWI objective and travel time misfit is limited in scope: in
the fifth section, I show that it does not hold in general for
transmitted wave data in which multiple rays connect source and
receiver (multiiple wave arrivals, non-simple metric).

@@

The parameter to be determined in the inverse problem studied here is
the bulk modulus $\kappa$. It is presumed is bounded and smooth:
\begin{equation}
  \label{eqn:coeff}
  \log \kappa \in C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)
\end{equation}
Under these hypotheses, the symmetric hyperbolic system \ref{eqn:awe}
has a unique distribution solution (for example \cite{Lax:PDENotes},
Theorem 7.4). The Propagation of
Singularities Theorem (\cite{Tay:81}, Theorem VI.2.1) %and
%the pull-back property of distributions (\cite{Dui:95}, Proposition
%1.3.3)
imples that $p(\cdot,\cdot;\bx_s)$ is smooth away from $\{\bx_s\}
\times \bR$, so that the traces $p(\bx_r, \cdot;\bx_s)$ on
$\{(\bx_r,t): t \in \bR\}$ are well-defined.

Define the {\em modeling operator}
\begin{equation}
  \label{eqn:modop}
  F: exp(M)  \rightarrow \Pi_{X_{sr}} C^{\infty}(\bR)
\end{equation}
by
\begin{equation}
  \label{eqn:modopdef}
  F[\kappa]= p|_{X_{sr} \times \bR}
\end{equation}

@@

The analysis to follow requires three further constraints on the set
of source-receiver location pairs $X_{sr}$ and the domain
$M \subset C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)$ of $F$. The first is
the {\em Exponential Decay} assumption:
\begin{quote}
  For any $n > 0$, $f \in C_0^{\infty}(\bR)$,
  there exist $K >
  0, T \in \bR,$ and $\delta>0$ so
  that for any $k$ with $0 \le k \le n$, $(\bx_s,\bx_r) \in
 X_{sr},$ and any $\log \kappa \in M$, the solution $(p,\bv)$ of \ref{eqn:awe} is smooth
  for $t>T$ and satisfies
  \[
    |\partial_t^k p(\bx_r,t;\bx_s)| \le K e^{-\delta t}, \, t > T.
  \]
\end{quote}
SInce $X_{sr}$ is bounded, the property described in this assumption
is a special case of {\em local energy decay}, a much studied topic. The {\em
  non-trapping} hypothesis is sufficient to ensure local energy decay:
all rays of geometric asymptotics (that is, geodesics for the Riemannian metric associated
to the system \ref{eqn:awe}) escape to infinity (\cite{Hristova:09},
Theorem 2, \cite{EgorovShubin}, Theorem 2.104, p. 192). That is,
exponential decay follows from a characteristic of $\kappa$.

The second and third assumptions explicitly involves a constraint on ray geometry
(therefore on $\kappa$). A {\em normal neighborhood} of
$\bx_s \in \bR^3$ is an open subset $\Omega$ of $\bR^3$ 
within which every point is connected to $\bx_s$ by  a unique ray  (\cite{Friedlander:75},
p. 16). Therefore, to every point $\bx \in \Omega$ is associated a
unique time of travel (geodesic distance) $\tau(\bx_s,\bx)$ from $\bx_s$ to $\bx$.
The {\em Single Arrival} assumption is
\begin{quote}
  $\Omega\subset \bR^3$ and $M \subset C^{\infty}(\bR^3) \cap
  L^{\infty}(\bR^3)$ are bounded open sets for which $X_{sr} \subset
  \Omega \times \Omega$, and for every source point $\bx_s\in X_{s}$
  and $\log \kappa \in M$, $\Omega$ is a normal neighborhood of
  $\bx_s$.
\end{quote}

The {\em No Return} assumption is
\begin{quote}
  for every source point $\bx_s \in X_{s}$, $\log \kappa \in M$, ray ${\bf X}:\bR^+\rightarrow \bR^3$
  with ${\bf X}(0)=\bx_s$ and any $t>0$, either $\{{\bf X}(s): 0\le s \le t\} \subset \Omega$
  or $\{{\bf X}(s): t < s < \infty\} \subset \bR^3 \setminus \Omega$.
\end{quote}
This assumption states that once a ray leaves
$\Omega$, its subsequent intersection with $\Omega$ is null (hence the name).

Thanks to Exponential Decay, each trace $p(\bx_r,\cdot;\bx_s)$ is in $L^{p}(\bR)$ for $p
\ge 1$.
Denote by $L^p(X_{sr} \times \bR)$ the product space $l^p(X_{sr},L^p(\bR))$
with the usuall product norm.

@@

AWI modifies the least-squares FWI problem by interposing a convolution operator or
{\em filter} $u \in L^2(X_{sr} \times \bR)$. That is, the residual
data $F[\kappa]-d$ is replaced by the filtered residual data $F[m]*u-d$.
Since $F[\kappa] \in L^1(X_{sr}\times \bR)$, the filtererd residual
makes sense.
The filter $u$ is intended to shift localized wave energy in $F[m]$ to
match that present in $d$. If on the other hand $u(\bx_r,t;\bx_s) =
\delta(t)$ for each $(\bx_x,\bx_r) \in X_{sr}$, then $tu = 0$. This
observation motivates the definition of the AWI objective function
\begin{equation}
  \label{eqn:awidefmain}
  J_{\rm AWI,\sigma}[\kappa;d] = \sum_{(\bx_s,\bx_r) \in X_{sr}}
  \frac{\|Tu_{\sigma}[\kappa;d](\bx_r,\cdot;\bx_s)\|^2}{\|u_{\sigma}[\kappa;d](\bx_r,\cdot;\bx_s)\|^2}.
\end{equation}
in which $Tu(\bx_r,t;\bx_s) = tu(\bx_r,t;\bx_s)$ and
\begin{equation}
  \label{eqn:reg}
  u_{\sigma}[\kappa;d] = \mbox{argmin}_{u \in L^2(X_{sr} \times \bR)} (\|F[\kappa]*u-d\|^2 + \sigma\|u\|^2)
\end{equation}
The objective function for the closely related MSWI approach,
mentioned in the introduction, is  
\begin{equation}
  \label{eqn:mswidefmain}
  J_{\rm MSWI,\sigma}[\kappa;d] = \sum_{(\bx_s,\bx_r) \in X_{sr}}
  \|Tu_{\sigma}[\kappa;d](\bx_r,\cdot;\bx_s)\|^2.
\end{equation}
Note that while the Lax-Milgram Theorem (\cite{Yosida}, section
III.7) ensures the existence of a unique solution
$u_{\sigma}[\kappa;d] \in L^2(X_{sr}\times \bR)$ of the least squares
problem \ref{eqn:reg},
it is not immediiately obvious that $Tu_{\sigma}[\kappa;d] \in L^2(X_{sr}\times \bR)$.

Asymptotic analysis of AWI and MSWI is based on a family of source
wavelets: for $\lambda > 0$,
\begin{equation}
  \fl(t)= \lambda^{\frac{1}{2}}f_1\left(\frac{t}{\lambda}\right).
\end{equation}
The relation between the dimensionless scale parameter $\lambda$ and
RMS wavelength will be explained below. Correspondingly, denote the
modeling operator with $f=\fl$ by $\Fl$.

@@

\begin{theorem}
  \label{thm:main}
  Assume that Exponential Decay, Single Arrival, and No Return
  hold for $X_{sr} \subset \Omega \times \Omega, M \subset C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)$.
  Select $r >0$ with units of
  pressure$^2 \times$time$^2$ and $\log \kappa^* \in M$. For $0 <
  \lambda $, set
  $\dl=\Fl[\kappa^*]$ (``noise-free data''), and  define $\ul[\kappa,\dl]$
  to be the solution of 
  equation \ref{eqn:reg} with  $\sigma=r\lambda$, $F=\Fl$ and $d=\dl$..
  There exist $0 < C_* \le C^*$, $0 < C$, and $W: M \times X_{sr}
  \rightarrow \bR^+$ with $C_* \le
  W \le C^*$ so that for
  $\log \kappa \in M$, $(\bx_s,\bx_r) \in X_{sr}$,
  \begin{itemize}
  \item[1. ] $Tu_{\sigma}[\kappa;\dl](\bx_r,\cdot;\bx_s) \in
    L^2(\bR)$;
  \item[2. ] for $\lambda > 0$,
    \begin{equation}
      \label{eqn:mswitt}
      |\|T\ul[\kappa;\dl](\bx_r,\cdot;\bx_s)\|- \lambda^{-\frac{1}{2}} (W[\kappa^*,\kappa](\bx_s,\bx_r))^{\frac{1}{2}}|\tau[\kappa](\bx_s,\bx_r)-\tau[\kappa^*](\bx_s,\bx_r)|| \le C
      \lambda^{\frac{1}{2}};
    \end{equation}
  \item[3. ]for sufficiently small $\lambda$, 
    \begin{equation}
      \label{eqn:awitt}
      \left|\frac{\|T\ul[\kappa;\dl](\bx_r,\cdot;\bx_s)\|}{\|\ul[\kappa,\dl](\bx_r,\cdot;\bx_s)\|}-
        |\tau[\kappa](\bx_s,\bx_r)-\tau[\kappa^*](\bx_s,\bx_r)|\right| \le C \lambda
    \end{equation}
  \end{itemize}
\end{theorem}

Summation over the source-receiver positions in $X_{sr}$ yields
\begin{cor}
  \label{thm:cor1}
  In addition to the assumptions of Theorem \ref{thm:main}, define $J_{\rm
    AWI,\sigma}$ and $J_{\rm MSWI,\sigma}$ by equations
  \ref{eqn:awidefmain}, \ref{eqn:mswidefmain} respectively. Then
\begin{equation}
\label{eqn:cor1mswi}   
\lim_{\lambda \rightarrow 0} \lambda J_{\rm MSWI,r\lambda}[\kappa;\dl]
= \sum_{(\bx_s,\bx_r) \in X_{sr}} W[\kappa^*,\kappa]^2(\bx_s,\bx_r) |\tau[\kappa](\bx_s,\bx_r) - \tau[\kappa^*](\bx_s,\bx_r)|^2.
\end{equation}
and
\begin{equation}
\label{eqn:cor1awi}   
\lim_{\lambda \rightarrow 0} J_{\rm AWI,r\lambda}[\kappa;\dl]
= \sum_{(\bx_s,\bx_r) \in X_{sr}}  |\tau[\kappa](\bx_s,\bx_r) - \tau[\kappa^*](\bx_s,\bx_r)|^2.
\end{equation}
\end{cor}

\section{Analysis of AWI and MSWI}
THis section is devoted to the proofs of the main result, Theorem
\ref{thm:main}.  The principal tool in the analysis is Hadamard's
decomposition of the acoustic Green's function, described in the
second
subsection below, along with some of its consequences, the chief being the
square-integrability of the time-scaled regularized filter
$tu_{\sigma}$. The next subsection reveals the link to travel time via
an approximate solution of the regularized least squares problem
\ref{eqn:reg}, based on the leading term in Hadamard[s decomposition.
The final subsection estimates the remainder and finishes the proof of
the main result.

\subsection{Hadamard's Decomposition}
The Hadamard decomposition of the acoustic Green's function \cite[]{Friedlander:75,Qian:JCP24} provides a
basis for analysis of AWI and MSWI. As presented by
\cite{Friedlander:75}, for instance, this construction applies to
second-order wave equations, but the system \ref{eqn:awe} is
equivalent fo a second-order problem: differentiate the first equation
with respect to $t$, eliminate $\bv$ via the second equation, and
rearrange to obtain
\begin{equation}
  \label{eqn:awe2}
  \frac{\rho}{\kappa}\frac{\partial^2 p}{\partial t^2} - \nabla^2 p -
  \nabla \log \rho \cdot \nabla p = \rho \frac{\partial f}{\partial t}
  \delta(\bx-\bx_s), \nonumber \\
  p=0, \, t \ll 0.
\end{equation}
If $(p,\bv)$ is a distribution solutions of the system
\ref{eqn:awe} then $p$ is a solution of equation
\ref{eqn:awe2}, as follows by differentiating the first equation of
\ref{eqn:awe} with respect to $t$ and eliminating $\bv$ via the second
equation. Conversely, if $p$ is a solution of the second-order initial
value problem\ref{eqn:awe2}, then under the conditions imposed in the
previous section, those to be imposed below),
\[
  \bv = \beta \int_{-\infty}^t p
\]
defines a distribution, and $(p,\bv)$ is a weak solutino of the system \ref{eqn:awe}.

Hadamard's construction is local, in both space-time and in the set of
acoustic models. In the present context, it requires the Single
Arrival hypothesis, and for global time validity the No Return
assumption.

\cite{Friedlander:75} produces for each $\log \kappa \in M$ functions $\tau[\kappa]$ (travel time), $a[\kappa]$
(geometric amplitude) on $X_s \times \Omega$ and $b[\kappa]$
(remainder kernel) on $X_s \times \Omega \times \bR$, smooth off the
diagonal - that is, for each $\bx_s \in X_s$, $\tau[\kappa](\bx_s,\cdot) \in C^{\infty}(\Omega \setminus
  \{\bx_s\})$, and similarly for $a[\kappa],b[\kappa]$; The traveltime $\tau$
  satisfies the eikonal equation
  \[
    \sqrt{\frac{\kappa}{\rho}} |\nabla \tau(\bx_s,\cdot)| = 1 \mbox{ in } \Omega \setminus
    \{\bx_s\}, \, \lim_{\bx \rightarrow \bx_s} \tau(\bx_s,\bx) = 0.
  \]
A by-product of the construction is a bound $A>0$ so that for any $\log \kappa \in M$
  and $(\bx_s,\bx_r) \in X_{sr}$,  $|\log a[\kappa](\bx_s,\bx_r)| <
  A$;  With no further assumptionss, there is also a uniform bound on $b(\bx_r,t;\bx_s)$ and
  its derivatives over any finite $t$-interval. However, with the
  addition of the Exponetial Decay assumption, the bound becomes
  global in time: for any $k \ge 0$, there exists $B \ge 0,\delta > 0$ so
  that for any $(\bx_s,\bx_r) \in X_{sr}$, $\log \kappa \in M$, 
\begin{equation}
  |\partial^k_t b[\kappa](\bx_r,t;\bx_r,s)| \le B e^{-\delta t}
  \label{eqn:best}
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%
